manuscript_figures

### 2024-06-27 scripts/data_cleaning.Rmd
updated HPI

#make QUAD data long
# Function to rename columns using metadata
# Get the unique column names from the QUAD ataframe
column_names.quad <- colnames(QUAD)

# Create a lookup table with name-to-code mappings
name_to_code.quad <- setNames(species.metadata$code, species.metadata$name)

@@

# use rbind.fill() method in R to combine data frames with different columns
QUAD_data <- rbind.fill(QUAD_long, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)

#remove 131-138 columns for observers not needed information
QUAD_data <- QUAD_data[ -c(14:17) ]

>>>>

#make QUAD data long
# Function to rename columns using metadata
# Get the unique column names from the QUAD dataframe
column_names.quad <- colnames(QUAD)
# Create a lookup table with name-to-code mappings
name_to_code.quad <- setNames(species.metadata$code, species.metadata$name)

@@

# Rename Kings Beach to King's Beach
quad_long_with_type$SITE[quad_long_with_type$SITE == "Kings Beach"] <- "King's Beach"

# use rbind.fill() method in R to combine data frames with different columns
QUAD_data <- rbind.fill(quad_long_with_type, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)

#remove 131-138 columns for observers not needed information
QUAD_data <- QUAD_data[ -c(14:19) ]

### 2024-06-27 scripts/manuscript_figures.Rmd
updated HPI

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(flextable)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Rename fort weatherill
fish.filt$SITE[fish.filt$SITE == "Kings Beach"] <- "King's Beach"

@@

Production to biomass ratios (P:B) for each fish species
```{r}
# load fish data from 2016
PB <- read.csv("data/fish.data.clean.calc.csv")

@@

>>>>

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(vegan)
library(flextable)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Rename kings beach
fish.filt$SITE[fish.filt$SITE == "Kings Beach"] <- "King's Beach"

@@

Production to biomass ratios (P:B) and habitat productivity index for each fish species
```{r}
# load fish data from 2016
PB <- read.csv("data/fish.data.clean.calc.csv")

@@

nMDS or PCA on benthic community structure
```{r}
# Load in QUAD data and filter for invertebrates
invert.filt <- QUAD %>%
  filter(type %in% c("molluscs", "echinoderms", "crustacean"))

# I noticed that 06/30/23 dates were accidentally labeled 06/30/29
# Replace "29" with "23" in the DATE column
invert.filt$DATE <- sub("29$", "23", invert.filt$DATE)

# Rename KB q to transect 6
invert.filt$TRANSECT[invert.filt$TRANSECT == "KB q"] <- "6"

# Sum count data at the transect level
invert.count <- invert.filt %>%
  group_by(TRANSECT, YEAR, SITE, type) %>%
  summarize(count.sum = sum(COUNT))

# Divide by 6 quadrats for correct abundance g/m2 calculation
invert.count$abundance <- invert.count$count.sum / 6

# Check for normality
shapiro.test(invert.count$abundance) #not normal so use nMDS

#make community matrix - extract columns with abundance information
com = invert.count[,6:ncol(invert.count)]

# Calculate Bray-Curtis dissimilarity
m_com <- vegdist(com, method = "bray")

# Check for warnings and errors
if (any(warnings())) {
  warning("Warnings during distance calculation:", warnings())
}

# Remove empty rows (if applicable)
m_com <- as.matrix(m_com[complete.cases(m_com), ])

# Check dimensions of m_com after removing empty rows
dim(m_com)

# Perform nMDS with Bray-Curtis dissimilarity
nmds <- metaMDS(m_com, distance = "bray")

# Create a dataframe for plotting
plot_data <- data.frame(
  NMDS1 = nmds$points[, 1],  # NMDS dimension 1
  NMDS2 = nmds$points[, 2],  # NMDS dimension 2
  type = invert.count$type,  # Type of invertebrate
  abundance = invert.count$abundance  # Abundance of invertebrate
)

# Plot nMDS with type labels and scaled points by abundance
ggplot(plot_data, aes(x = NMDS1, y = NMDS2, color = type, size = abundance)) +
  geom_point(alpha = 0.8) +
  scale_size_continuous(range = c(2, 8)) +  # Adjust point size range as needed
  labs(title = "nMDS Plot of Invertebrate Types by Abundance",
       x = "NMDS 1", y = "NMDS 2") +
  theme_minimal() +
  theme(legend.position = "right") +
  geom_text(aes(label = type), nudge_x = 0.05, nudge_y = 0.05, size = 3)  # Add type labels
```
